More equivalences for Open pNets\thanks{This work was partially funded by the Associated Team FM4CPS between INRIA and ECNU, Shanghai}

### 2017-02-10 2016-Weak-FH-Bisim/WeakBisim.tex
transitivity V2 !

       It is trivial to check reflexivity and symmetry. Here we focus on the
transitivity. 
To prove transitivity of strong FH-bisimulation on pNets it is sufficient to prove 
transitivity of the strong FH-bisimulation on states. Consider 3 open automata 
$\mathcal{T}_1$, $\mathcal{T}_2$, $\mathcal{T}_3$ and states $s^1$, $s^2$, $s^3$ in those 
automata\footnote{We omit the constraints stating that each $s^i_x$ is in the states of 
$\mathcal{T}_i$ for the sake of readability}.
Suppose we have $\mathcal{R}$ an FH-bisimulation relation between states of 
$\mathcal{T}_1$ and of  $\mathcal{T}_2$; members of $\mathcal{R}$ are of the form 
$(s^1,s^2|\Pred)$.
Suppose we also  have $\mathcal{R}'$ an FH-bisimulation relation between states of 
$\mathcal{T}_2$ and of  $\mathcal{T}_3$; members of $\mathcal{R}2$ are of the form 
$(s^2,s^3|\Pred\,')$.

Let $\mathcal{R}''$ be the relation: 
\[\mathcal{R}'' = \{(s^1,s^3|\Pred\,'')|\,\Pred\,''=\bigvee_{ \begin{array}{l}
(s^1,s^2|\Pred)\in\mathcal{R}\\ (s^2,s^3|\Pred\,')\in\mathcal{R}' 	
\end{array}
}\,\Pred\land\Pred\,'\}\]

The relation is built as follows: for each pair of state $s^1$, $s^3$, for each state 
$s^2$ such that $\mathcal{R}$ relates $s^1$ and $s^2$, and $\mathcal{R}'$ relates $s^2$ 
and $s^3$, we take the conjunction of the two predicates. The predicates for different 
values of $s^2$ are collected by a disjunction.

 We will show
that $\mathcal{R}''$ is an FH-bisimulation. Consider 
$(s^1,s^3|\Pred\,'')\in\mathcal{R}''$. Then there is a set of states of 
$\mathcal{T}_2$ relating $s^1$ and $s^3$, let $(s^2_p)^{p\in P}$ be this family.

For any $p\in P$ by definition of $\mathcal{R}''$,
$(s^1,s^2_p|\Pred_p)\in\mathcal{R}$,  and $(s^2_p,s^3|\Pred\,'_p)\in\mathcal{R}$. We have 
the 
following by definition of bisimulation.
For any open transition $OT$ in $\mathcal{T}_1$ originating from $s^1$:
Additionally we have $\Pred\,'' = \bigvee_{p\in P} \Pred_p\land\Pred_p\,'$.
\begin{mathpar}

@@

\end{mathpar}
such that  $\forall y, J_{2px}=J_{3pxy}, 
(s^{2}_x,s^{3}_{pxy}|\Pred_{target_{pxy}})\in \mathcal{R}'$; and  \\
$\Pred\,'_p \land \Pred_{OT_x}\\
\hspace{1cm} \implies\!\!\! \bigvee_{y\in Y}
\left( \forall j. b_{jpx}=b_{jpxy}  \Rightarrow \Pred_{OT_{pxy}}
\land v_{px}\!=\!v_{pxy} \land
\Pred_{target_{pxy}}\subst{\Post_{OT_{px}}}\subst{\Post_{OT_{pxy}}}\right)$.

@@

First, $\forall y, \forall x, \forall p, J_1=J_{2px}=J_{3pxy}, 
({s^{1}}',s^{3}_{pxy}|\Pred\,'_{target_{pxy}})\in \mathcal{R}''$. 
Indeed for any 
$p$, 
$x$, and 
$y$, $s^2_{px}$
relate ${s^{1}}'$ and $s^{3}_{pxy}$, indeed
$(s^{1'},s^{2}_{px}|\Pred_{target_{px}})\in \mathcal{R}$
 and $(s^{2}_x,s^{3}_{pxy}|\Pred_{target_{pxy}})\in \mathcal{R}'$. 
 More precisely, $s^{2}_{px} \in ({s^{2}_p}')^{p\in P'}$ where $({s^{2}_p}')^{p\in 
 P'}$ 
 is 
 the set of states relating ${s^{1}}'$ and $s^{3}_{pxy}$.
Consequently, for all $p$, $x$, $y$, $\Pred_{target_{px}}\land 
\Pred_{target_{pxy}}\implies 
\Pred\,'_{target_{pxy}}$ (this is one element of the  disjunction defining the 
predicate 
relating ${s^{1}}'$ and $s^{3}_{pxy}$ in the definition of $\mathcal{R}''$).

@@

Additionally, the Post renamings do not modify the predicates that only use non-renamed 
variables, and because of the domain of the renamings:\\
$\Pred_{target_{px}}\subst{\Post_{OT}}\subst{\Post_{OT_{px}}}\land 
\Pred_{target_{pxy}}\subst{\Post_{OT_{px}}}\subst{\Post_{OT_{pxy}}}\\
= 
\Pred_{target_{px}}\subst{\Post_{OT}}\subst{\Post_{OT_{px}}}\subst{\Post_{OT_{pxy}}}\land 
\Pred_{target_{pxy}}\subst{\Post_{OT}}\subst{\Post_{OT_{px}}}\subst{\Post_{OT_{pxy}}}\\
\implies\Pred\,'_{target_{pxy}}\subst{\Post_{OT}}\subst{\Post_{OT_{px}}}\subst{\Post_{OT_{pxy}}}\\
= \Pred\,'_{target_{pxy}}\subst{\Post_{OT}}\subst{\Post_{OT_{pxy}}} $

This allows us to conclude: using $\Pred\,'' = \bigvee_{p\in P} \Pred_p\land\Pred_p\,'$ 
and the assertions above we obtained the needed property.

>>>>

       It is trivial to check reflexivity and symmetry. Here we focus on the
transitivity. 
To prove transitivity of strong FH-bisimulation on pNets it is sufficient to prove 
transitivity of the strong FH-bisimulation on states. Consider 3 open automata 
$\mathcal{T}_1$, $\mathcal{T}_2$, $\mathcal{T}_3$ and states $s^1$, $s^2$, $s^3$ in those 
automata\footnote{We omit the constraints stating that each $s^i_x$ is in the states of 
$\mathcal{T}_i$ for the sake of readability}.
Suppose we have $\mathcal{R}$ an FH-bisimulation relation between states of 
$\mathcal{T}_1$ and of  $\mathcal{T}_2$; members of $\mathcal{R}$ are of the form 
$(s^1,s^2|\Pred)$.
Suppose we also  have $\mathcal{R}'$ an FH-bisimulation relation between states of 
$\mathcal{T}_2$ and of  $\mathcal{T}_3$; members of $\mathcal{R}2$ are of the form 
$(s^2,s^3|\Pred\,')$.
We use $\fv$ a function returning the free variables of a term/automaton.

Let $\mathcal{R}''$ be the relation: 
\[\mathcal{R}'' = \Big\{(s^1,s^3|\Pred\,''\phi)\Big|
\displaystyle\Pred\,''=\bigvee_{ \begin{array}{l}
(s^1,s^2|\Pred)\in\mathcal{R}\\ (s^2,s^3|\Pred\,')\in\mathcal{R}' 	
\end{array}
}\!\!\!\!\!\exists \phi.\begin{array}[t]{l} (\Pred\land\Pred\,')\phi\\
\land \fv((\Pred\land\Pred)\phi)\subseteq \fv(\mathcal{T}_1)\cup\fv(\mathcal{T}_3)
\\
\land \dom(\phi)\subseteq \fv(\mathcal{T}_2)
\end{array}\Big\}\]

\TODO{is it satisfactory as a notion of transitive composition? It would be easy to 
extend it a bit to allow having several substitutions for the same s2, eg one that works 
for different values of variables in T1 and T3. It would make the notations more complex 
but it would be OK. I do not think I can do more than that}

The relation is built as follows: for each pair of state $s^1$, $s^3$, for each state 
$s^2$ such that $\mathcal{R}$ relates $s^1$ and $s^2$, and $\mathcal{R}'$ relates $s^2$ 
and $s^3$, we take the conjunction of the two predicates. The predicates for different 
values of $s^2$ are collected by a disjunction. 
Finally, a substitution should ensure that the 
free variables of the predicates are only the ones of $\mathcal{T}_1$ and 
$\mathcal{T}_1$. This substitution somehow matches the two predicates by eliminating the 
free variables.

 We will show
that $\mathcal{R}''$ is an FH-bisimulation. Consider 
$(s^1,s^3|\Pred\,'')\in\mathcal{R}''$. Then there is a set of states of 
$\mathcal{T}_2$ relating $s^1$ and $s^3$, let $(s^2_p)^{p\in P}$ be this family.  
For any $p\in P$ by definition of $\mathcal{R}''$,
$(s^1,s^2_p|\Pred_p)\in\mathcal{R}$,  and $(s^2_p,s^3|\Pred\,'_p)\in\mathcal{R}$. 
Additionally we have $\Pred\,'' = (\bigvee_{p\in P} \Pred_p\land\Pred_p\,'\phi_p)$.
We have 
the 
following by definition of bisimulation.
For any open transition $OT$ in $\mathcal{T}_1$ originating from $s^1$:

@@

\end{mathpar}
such that  $\forall y, J_{2px}=J_{3pxy}, 
(s^{2}_x,s^{3}_{pxy}|\Pred_{target_{pxy}})\in \mathcal{R}'$; and  \\
$\Pred\,'_p \land \Pred_{OT_{px}}\\
\hspace{1cm} \implies\!\!\! \bigvee_{y\in Y}
\left( \forall j. b_{jpx}=b_{jpxy}  \Rightarrow \Pred_{OT_{pxy}}
\land v_{px}\!=\!v_{pxy} \land
\Pred_{target_{pxy}}\subst{\Post_{OT_{px}}}\subst{\Post_{OT_{pxy}}}\right)$.

@@

First, $\forall y, \forall x, \forall p, J_1=J_{2px}=J_{3pxy}, 
({s^{1}}',s^{3}_{pxy}|\Pred\,'_{target_{pxy}})\in \mathcal{R}''$. 
Indeed for any 
$p$, 
$x$, and 
$y$, $s^2_{px}$
relate ${s^{1}}'$ and $s^{3}_{pxy}$, indeed
$(s^{1'},s^{2}_{px}|\Pred_{target_{px}})\in \mathcal{R}$
 and $(s^{2}_x,s^{3}_{pxy}|\Pred_{target_{pxy}})\in \mathcal{R}'$. 
 More precisely, $s^{2}_{px} \in ({s^{2}_p}')^{p\in P'}$ where $({s^{2}_p}')^{p\in 
 P'}$ 
 is 
 the set of states relating ${s^{1}}'$ and $s^{3}_{pxy}$.
Consequently, for all $p$, $x$, $y$,\\
 $\left(\exists \phi'.\,(\Pred_{target_{px}}\land 
\Pred_{target_{pxy}})\phi'\land \fv((\Pred\land\Pred)\phi')\subseteq 
\fv(\mathcal{T}_1)\cup\fv(\mathcal{T}_3)\right)\implies 
\Pred\,'_{target_{pxy}}$ \\
This is one element of the  disjunction defining the 
predicate 
relating ${s^{1}}'$ and $s^{3}_{pxy}$ in the definition of $\mathcal{R}''$.

@@

Additionally, the Post substitutions only have an effect on the  predicates that  use 
some of the 
substituted 
variables, and because of the domain of the substitutions we have:\\
$(\Pred_{target_{px}}\subst{\Post_{OT}}\subst{\Post_{OT_{px}}}\land 
\Pred_{target_{pxy}}\subst{\Post_{OT_{px}}}\subst{\Post_{OT_{pxy}}})\phi\\
%
= 
(\Pred_{target_{px}}\subst{\Post_{OT}}\subst{\Post_{OT_{px}}}\subst{\Post_{OT_{pxy}}}\land
\Pred_{target_{pxy}}\subst{\Post_{OT}}\subst{\Post_{OT_{px}}}\subst{\Post_{OT_{pxy}}}) 
\phi\\
%
= 
((\Pred_{target_{px}}\land
\Pred_{target_{pxy}})\subst{\Post_{OT_{px}}}\phi) 
\subst{\Post_{OT}}\subst{\Post_{OT_{pxy}}}\\
%
\implies\Pred\,'_{target_{pxy}}\subst{\Post_{OT}}
\subst{\Post_{OT_{pxy}}}\\
 $
Finally, gathering the previous results, and using the fact that only variables of 
$\mathcal{T}_2$ are changed by $\phi$ substitutions:

$\Pred\,'' \land \Pred_{OT}\\
\hspace{1cm} \implies\!\!\! \bigvee_{p\in P}(\Pred_p \land 
\Pred_{OT}\land\Pred\,'_p)\phi\\
\hspace{1cm} \implies\!\!\! \bigvee_{x\in X}\bigvee_{y\in Y}\bigvee_{p\in P}
\left( \forall j. b_j=b_{jpxy}  \Rightarrow \Pred_{OT_{pxy}}
\land v\!=\!v_{pxy} \land
\Pred\,'_{target_{pxy}}\subst{\Post_{OT}}\subst{\Post_{OT_{pxy}}}\right)
$

This is the expected property.
